A Survey of Real-Time Strategy Game AI\\ Research and Competition in StarCraft

### 2013-08-09 survey.tex
Additional proofreading

\subsubsection{Task Decomposition}
For all the previous reasons, most existing approaches to play games as StarCraft work by decomposing the problem of playing an RTS game into a collection of smaller problems, to be solved independently. Specifically, a common subdivision is:
\begin{itemize}
\item {\em Strategy}: corresponds to the high-level decision making process. This is
 the highest level of abstraction for the game comprehension. Finding an efficient strategy or
 counter-strategy against a given opponent is key in RTS games. It concerns the whole
 set of units and buildings a player owns. %If we may allow ourselves a military analogy, strategy is orders given by generals at the HQ. 
\item {\em Tactics}: are the implementation of the current strategy. It
 implies army and building positioning, movements, timing, and so
 on. Tactics concerns a group of units. %They are orders given by officers on the battlefield.
\item {\em Reactive control}: is the implementation of tactics. This
 consists in moving, targeting, firing, fleeing, hit-and-run
 techniques (also knows as ``kiting'') during battle. Reactive control focuses
 on a specific unit. %It is soldiers applying officers' orders. 
%\end{itemize}
%
%\noindent
%Both  strategy  and  tactics  are  usually  called
%\textit{macro-management}, whereas reactive control is also called
%\textit{micro-management}. In addition of the three tasks shown above,
%one could also consider these two tasks that an AI has to deal with:
%\begin{itemize}
\item {\em Terrain analysis}: consists in the analysis of regions composing
 the map: choke-points, minerals and gas emplacements, low and high
 walkable grounds, islands, etc.
\item {\em Intelligence gathering}: corresponds to information collected about
 the opponent. Because of the fog-of-war, players must
 regularly send scouts to localize and spy enemy bases.%, and must see or infer what buildings and technology the enemy has developed in order to adapt its strategy.
\end{itemize}

@@

Holistic approaches to address RTS AI attempt to address the whole problem using a single unified method. To the best of our knowledge, with a few exceptions, such as the Darmok system \cite{OntanonMSR10} (which uses a combination of case-based reasoning and learning from demonstration) or ALisp \cite{Marthi05}, there has not been much work in this direction.   The main reason is that the complexity of RTS games is too large, and approaches that decompose the problem into smaller, separate, problems, achieve better results in practice. However, holistic approaches, based, for example, on Monte Carlo Tree Search, have only been explored in the context of smaller-scale RTS games~\cite{ontanon2013naive}, but techniques that scale up to large RTS games as StarCraft are still not available.

@@

Nevertheless, from the observations we made during the tournament,
we can draw the conclusion that the available bots are still 
very constrained. No bot in the competition played the Zerg race,
which is surprising as the AIIDE 2010 winner (Overmind) did so.
Presumably, implementing a good Zerg strategy is more demanding
than for the Protoss or Terran races. Many bots consistently crashed
when playing against a random race built-in bot for testing, and
also did so when the map size was changed from $128\times 128$ to 
any other. Furthermore, every single bot sometimes failed to finish
off an already beaten opponent, such that the game had to be stopped
after a previously determined maximum time. It also seems that most of
the current bots are not very good at adapting their strategy to the
one of their opponent during a game, or at least (via the read/write
procedure of game information) within a series of games.

@@

One of the major conclusions from the results of the StarCraft competitions is the lack of adaptation of bots.  Some switch between different build-orders, but do not fully adapt their strategy. No bot is capable of observing the opponent and autonomously synthesize a good plan from scratch to counter the opponent strategy. In this section we analyzed this claim quantitatively using the tools of \cite{SynnaeveOpeningCig11}. We analyzed the replays from the 2011 and 2012 AIIDE competitions (shown in Figures \ref{fig:aiide2011-botopenings} and \ref{fig:aiide2012-botopenings} respectively). We analyzed the way bots choose their openings depending on which other bot they are playing against. Given that there is no dominant strategy in StarCraft, and it is necessary to see what the opponent is doing in order to determine the best opening, we would expect bots to change their openings depending on which other bot they are playing (since each bot uses a different strategy, or set of strategies). Using clustering, we identified the most common openings in all the replays from the 2011 and 2012 competitions (each one shown with a different color in the figures). No data is shown for the ItayUnvermind bot, since its opening did not match significantly with any of the ones used in our study (extracted from humans pro-gamers).

>>>>

\subsubsection{Task Decomposition}
For all the previous reasons, most existing approaches to play games as StarCraft work by decomposing the problem of playing an RTS game into a collection of smaller problems, to be solved independently. Specifically, a common subdivision is:
\begin{itemize}
\item {\em Strategy}: corresponds to the high-level decision making process. This is
 the highest level of abstraction for the game comprehension. Finding an efficient strategy or
 counter-strategy against a given opponent is key in RTS games, and concerns the whole
 set of units a player owns. %If we may allow ourselves a military analogy, strategy is orders given by generals at the HQ. 
\item {\em Tactics}: are the implementation of the current strategy. It
 implies army and building positioning, movements, timing, and so
 on. Tactics concerns a group of units. %They are orders given by officers on the battlefield.
\item {\em Reactive control}: is the implementation of tactics. This
 consists in moving, targeting, firing, fleeing, hit-and-run
 techniques (also knows as ``kiting'') during battle. Reactive control focuses
 on a specific unit. %It is soldiers applying officers' orders. 
%\end{itemize}
%
%\noindent
%Both  strategy  and  tactics  are  usually  called
%\textit{macro-management}, whereas reactive control is also called
%\textit{micro-management}. In addition of the three tasks shown above,
%one could also consider these two tasks that an AI has to deal with:
%\begin{itemize}
\item {\em Terrain analysis}: consists in the analysis of regions composing
 the map: choke-points, minerals and gas emplacements, low and high
 walkable grounds, islands, etc.
\item {\em Intelligence gathering}: corresponds to information collected about
 the opponent. Because of the fog-of-war, players must
 regularly send scouts to localize and spy enemy bases.%, and must see or infer what buildings and technology the enemy has developed in order to adapt its strategy.
\end{itemize}

@@

Holistic approaches to address RTS AI attempt to address the whole problem using a single unified method. To the best of our knowledge, with a few exceptions, such as the Darmok system \cite{OntanonMSR10} (which uses a combination of case-based reasoning and learning from demonstration) or ALisp \cite{Marthi05}, there has not been much work in this direction.   The main reason is that the complexity of RTS games is too large, and approaches that decompose the problem into smaller, separate, problems, achieve better results in practice. However, holistic approaches, based, for example, on Monte Carlo Tree Search, have only been explored in the context of smaller-scale RTS games~\cite{ontanon2013naive}. Techniques that scale up to large RTS games as StarCraft are still not available.

@@

Nevertheless, from the observations we made during the tournament,
we can draw the conclusion that the available bots are still 
very constrained. No bot in the competition played the Zerg race,
which is surprising as the AIIDE 2010 winner (Overmind) did so.
Presumably, implementing a good Zerg strategy is more demanding
than implementing one for the Protoss or Terran races. Many bots consistently crashed
when playing against a random race built-in bot for testing, and
also did so when the map size was changed from $128\times 128$ to 
any other. Furthermore, every single bot sometimes failed to finish
off an already beaten opponent, such that the game had to be stopped
after a previously determined maximum time. It also seems that most of
the current bots are not very good at adapting their strategy to the
one of their opponent during a game, or at least (via the read/write
procedure of game information) within a series of games.

@@

One of the major conclusions from the results of the StarCraft competitions is the lack of adaptation of bots.  Some switch between different build-orders, but do not fully adapt their strategy. No bot is capable of observing the opponent and autonomously synthesize a good plan from scratch to counter the opponent strategy. In this section we analyzed this claim quantitatively using the tools of \cite{SynnaeveOpeningCig11}. We analyzed the replays from the 2011 and 2012 AIIDE competitions (shown in Figures \ref{fig:aiide2011-botopenings} and \ref{fig:aiide2012-botopenings} respectively). We analyzed the way bots choose their openings depending on which other bot they are playing against. Given that there is no dominant strategy in StarCraft, and that it is necessary to see what the opponent is doing in order to determine the best opening, we would expect bots to change their openings depending on which other bot they are playing (since each bot uses a different strategy, or set of strategies). Using clustering, we identified the most common openings in all the replays from the 2011 and 2012 competitions (each one shown with a different color in the figures). No data is shown for the ItayUnvermind bot, since its opening did not match significantly with any of the ones used in our study (extracted from humans pro-gamers).
